---
title: "2024_1_26_Female_Male_DEA_Tumor_TumorAdj"
output: html_document
date: "2024-01-26"
---

**Purpose:**

* Identify female and male tumor/tumor-adjacent differentially expressed genes with only library type as a covariate

**Methods:**

* Limma voom pipeline on all tumor-tumor-adjcent comparison

* Only covariate is library type

**Results:**

## Load libraries

```{r, message=FALSE, warning=FALSE}
library(edgeR)
library(DESeq2)
library(limma)
library(ggplot2)
library(dplyr)
```

## Read in data

```{r}
metadata <- read.table("~/3.0 Hasting Research/Liver Cancer Analysis/Japan_HCC_metadata_QC_filtered.csv", row.names=1,header=TRUE, sep=",")
tumorAdjacentExp <- read.csv("~/3.0 Hasting Research/Liver Cancer Analysis/Japan_HCC_Salmon_Expression_QC_filtered.csv", row.names = 1, header=TRUE)
colnames(tumorAdjacentExp) <- gsub("\\.", "-", colnames(tumorAdjacentExp))
identical(metadata$sampleid, colnames(tumorAdjacentExp))

table(metadata$gender_tissue_viral)
```
## Relabeling genes
```{r}
# Importing gene annotations
genes <- read.table("~/3.0 Hasting Research/Liver Cancer Analysis/gencodeTranscripts.txt", header=TRUE, sep="\t")
genes <- data.frame(genes)
tumorAdjacentExp <- tumorAdjacentExp[rownames(tumorAdjacentExp) %in% genes$GENEID ,]
genes <- genes[match(rownames(tumorAdjacentExp), genes$GENEID),]
identical(rownames(tumorAdjacentExp), genes$GENEID)
# Calculating gene length, this is needed for calculating the FPKM values
genes$length <- with(genes, end - start)
```

## Create DEG list object
```{r}
dge <- DGEList(counts=tumorAdjacentExp, genes=genes)

colnames(dge) <- colnames(tumorAdjacentExp)
dge$samples$ID <- metadata$ID
dge$samples$tumor <- factor(metadata$tumor)
dge$samples$library_type <- metadata$libtype
dge$samples$viral <- metadata$Virus_infection
dge$samples$gender_tussue <- metadata$gender_tissue
dge$samples$gender_tissue_viral <- metadata$gender_tissue_viral
```

## Filtering by FPKM expression level
```{r}
# Keeping genes that have a mean FPKM of at least 0.5 in at least one of the
# groups under investigation and at least 6 reads in at least 10 samples
fpkm <- rpkm(dge, gene.length=dge$genes$length)

M_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HBV")],1,mean,na.rm=TRUE)
M_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HBV")],1,mean,na.rm=TRUE)
M_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HCV")],1,mean,na.rm=TRUE)
M_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HCV")],1,mean,na.rm=TRUE)

F_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HBV")],1,mean,na.rm=TRUE)
F_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HBV")],1,mean,na.rm=TRUE)
F_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HCV")],1,mean,na.rm=TRUE)
F_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HCV")],1,mean,na.rm=TRUE)

keep <- (M_1_HBV_mean_fpkm > 0.5 | M_0_HBV_mean_fpkm > 0.5 | 
           M_1_HCV_mean_fpkm > 0.5 | M_0_HCV_mean_fpkm > 0.5 |
           F_1_HBV_mean_fpkm > 0.5 | F_0_HBV_mean_fpkm > 0.5 |
           F_1_HCV_mean_fpkm > 0.5 | F_0_HCV_mean_fpkm > 0.5 )

dge <- dge[keep,,keep.lib.sizes=FALSE]
dge <- calcNormFactors(dge, method="TMM")
keep <- rowSums(dge$counts > 6) >= 10
dge <- dge[keep,,keep.lib.size=FALSE]
dge <- calcNormFactors(dge, method="TMM")

# N of genes retained after filtering
dim(dge$genes)
```

## Create design matrix
```{r}
design <- model.matrix(~0+dge$samples$gender_tussue+dge$samples$library_type)
colnames(design) <- gsub("dge\\$samples\\$gender_tussue", "", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$library_typeunstranded", "library_type", colnames(design))
head(design)
```

## Voom analysis
```{r}
v <- voomWithQualityWeights(dge, design, plot=TRUE)
```

# DEA
```{r}
# Block design for individual. This is used in tumor-normal comparisons with
# paired samples.
corfit <- duplicateCorrelation(v, design, block = v$targets$ID)
# This should give a positive correlation value. It represents the
# correlation between measurements made on the same person.
corfit$consensus

# Fitting the linear model with limma.
# If using paired samples, the within-patient correlation and a block design
# for patient is used to account for pairwise samples
fit <- lmFit(v, design, block = v$targets$ID, correlation = corfit$consensus)

# Contrast design for differential expression
# Defining pairwise comparisons
contrasts <- makeContrasts(`F_1` - `F_0`,
                           `M_1` - `M_0`,
                           levels=colnames(design))
head(contrasts)

# Running contrast analysis
vfit <- contrasts.fit(fit, contrasts = contrasts)

# Computing differential expression based on the empirical Bayes moderation of
# the standard errors towards a common value. Robust = should the estimation of
# the empirical Bayes prior parameters be robustified against outlier sample
# variances?
veBayesFit <- eBayes(vfit, robust=TRUE)
# Look at N of DEGs with adj. p <0.01 and log2FC>2
summary(decideTests(veBayesFit, adjust.method = "BH", p.value = 0.05, lfc = 2))
plotSA(veBayesFit, main = "Final model: Mean-variance trend")

vTopTable_female <- topTable(veBayesFit, n=Inf, p.value=1, lfc=0, coef = 1)
write.csv(vTopTable_female, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/female_tumor_tumor_adj_p1_lfc0.csv")

DEGs_female <- topTable(veBayesFit, n=Inf, p.value=0.05, lfc=2, coef=1)
DEGs_female_print <- data.frame(DEGs_female$GENEID, DEGs_female$gene_name, DEGs_female$adj.P.Val, DEGs_female$logFC)
write.csv(DEGs_female_print, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/female_tumor_tumor_adj_p0.05_lfc2.csv")

vTopTable_male <- topTable(veBayesFit, n=Inf, p.value=1, lfc=0, coef = 2)
write.csv(vTopTable_male, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/male_tumor_tumor_adj_p1_lfc0.csv")

DEGs_male <- topTable(veBayesFit, n=Inf, p.value=0.05, lfc=2, coef=2)
DEGs_male_print <- data.frame(DEGs_male$GENEID, DEGs_male$gene_name, DEGs_male$adj.P.Val, DEGs_male$logFC)
write.csv(DEGs_male_print, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/male_tumor_tumor_adj_p0.05_lfc2.csv")
```

**Interpretation:**

* Model fit looks good before and after voom adjustment

* Positive correlation for samples from the same patient

* DEG lists saved

**Future steps:**

* Complete for male and female segregated and HBV/HCV segregated and then move on the visualization and downstream analysis
