---
title: "2024_4_15_Viral_sex_combined_DEA_down-sampling_males"
output: html_document
date: "2024-04-15"
---

**Purpose:**

* Identify viral and sex-stratified tumor/tumor-adjacent differentially expressed genes with only library type as a covariate

* Down-sample males to see if there are still more male-DEAs identified

**Methods:**

* Limma voom pipeline on all tumor-tumor-adjcent comparison

* Only covariate is library type

**Results:**

## Load libraries

```{r, message=FALSE, warning=FALSE}
library(edgeR)
library(DESeq2)
library(limma)
library(ggplot2)
library(dplyr)
```

## Read in data and down-sample

```{r}
metadata <- read.table("~/3.0 Hasting Research/Liver Cancer Analysis/Japan_HCC_metadata_QC_filtered.csv", row.names=1,header=TRUE, sep=",")
tumorAdjacentExp <- read.csv("~/3.0 Hasting Research/Liver Cancer Analysis/Japan_HCC_Salmon_Expression_QC_filtered.csv", row.names = 1, header=TRUE)
colnames(tumorAdjacentExp) <- gsub("\\.", "-", colnames(tumorAdjacentExp))
identical(metadata$sampleid, colnames(tumorAdjacentExp))

table(metadata$gender_tissue_viral)

F_HBV_adj <- metadata[which(metadata$gender_tissue_viral=="F_0_HBV"),]
F_HBV_tum <- metadata[which(metadata$gender_tissue_viral=="F_1_HBV"),]
F_HBV_tum <- F_HBV_tum[which(F_HBV_tum$ID %in% F_HBV_adj$ID),]

M_HBV_adj <- metadata[which(metadata$gender_tissue_viral=="M_0_HBV"),]
M_HBV_tum <- metadata[which(metadata$gender_tissue_viral=="M_1_HBV"),]
set.seed(1)
random_ds <- sample(M_HBV_adj$ID, 7)
M_HBV_adj <- M_HBV_adj[which(M_HBV_adj$ID %in% random_ds),]
M_HBV_tum <- M_HBV_tum[which(M_HBV_tum$ID %in% random_ds),]

F_HCV_adj <- metadata[which(metadata$gender_tissue_viral=="F_0_HCV"),]
F_HCV_tum <- metadata[which(metadata$gender_tissue_viral=="F_1_HCV"),]

M_HCV_adj <- metadata[which(metadata$gender_tissue_viral=="M_0_HCV"),]
M_HCV_tum <- metadata[which(metadata$gender_tissue_viral=="M_1_HCV"),]
M_HCV_adj <- M_HCV_adj[which(M_HCV_adj$ID %in% M_HCV_tum$ID),]
random_ds <- sample(M_HCV_adj$ID, 32)
M_HCV_adj <- M_HCV_adj[which(M_HCV_adj$ID %in% random_ds),]
M_HCV_tum <- M_HCV_tum[which(M_HCV_tum$ID %in% random_ds),]

metadata <- rbind(F_HBV_adj, F_HBV_tum, M_HBV_adj, M_HBV_tum, F_HCV_adj, F_HCV_tum, M_HCV_adj, M_HCV_tum)
table(metadata$gender_tissue_viral)
```

```{r}
tumorAdjacentExp <- tumorAdjacentExp[,match(metadata$sampleid, colnames(tumorAdjacentExp))]
identical(metadata$sampleid, colnames(tumorAdjacentExp))
```

## Validate that ages are not different

```{r}
t.test(F_HBV_tum$Age, M_HBV_tum$Age)
t.test(F_HBV_tum$Ta, M_HBV_tum$Ta)

t.test(F_HCV_tum$Age, M_HCV_tum$Age)
t.test(F_HCV_tum$Ta, M_HCV_tum$Ta)
```

## Relabeling genes
```{r}
# Importing gene annotations
genes <- read.table("~/3.0 Hasting Research/Liver Cancer Analysis/gencodeTranscripts.txt", header=TRUE, sep="\t")
genes <- data.frame(genes)


tumorAdjacentExp <- tumorAdjacentExp[rownames(tumorAdjacentExp) %in% genes$GENEID ,]
genes <- genes[match(rownames(tumorAdjacentExp), genes$GENEID),]
identical(rownames(tumorAdjacentExp), genes$GENEID)
# Calculating gene length, this is needed for calculating the FPKM values
genes$length <- with(genes, end - start)
```

## Create DEG list object
```{r}
dge <- DGEList(counts=tumorAdjacentExp, genes=genes)

colnames(dge) <- colnames(tumorAdjacentExp)
dge$samples$ID <- metadata$ID
dge$samples$tumor <- factor(metadata$tumor)
dge$samples$library_type <- metadata$libtype
dge$samples$viral <- metadata$Virus_infection
dge$samples$tissue_viral <- paste(metadata$Virus_infection, metadata$tumor, sep="_")
dge$samples$gender_tussue <- metadata$gender_tissue
dge$samples$gender_tissue_viral <- metadata$gender_tissue_viral
```

## Filtering by FPKM expression level
```{r}
# Keeping genes that have a mean FPKM of at least 0.5 in at least one of the
# groups under investigation and at least 6 reads in at least 10 samples
fpkm <- rpkm(dge, gene.length=dge$genes$length)

M_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HBV")],1,mean,na.rm=TRUE)
M_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HBV")],1,mean,na.rm=TRUE)
M_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HCV")],1,mean,na.rm=TRUE)
M_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HCV")],1,mean,na.rm=TRUE)

F_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HBV")],1,mean,na.rm=TRUE)
F_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HBV")],1,mean,na.rm=TRUE)
F_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HCV")],1,mean,na.rm=TRUE)
F_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HCV")],1,mean,na.rm=TRUE)

keep <- (M_1_HBV_mean_fpkm > 0.5 | M_0_HBV_mean_fpkm > 0.5 | 
           M_1_HCV_mean_fpkm > 0.5 | M_0_HCV_mean_fpkm > 0.5 |
           F_1_HBV_mean_fpkm > 0.5 | F_0_HBV_mean_fpkm > 0.5 |
           F_1_HCV_mean_fpkm > 0.5 | F_0_HCV_mean_fpkm > 0.5 )

dge <- dge[keep,,keep.lib.sizes=FALSE]
dge <- calcNormFactors(dge, method="TMM")
keep <- rowSums(dge$counts > 6) >= 10
dge <- dge[keep,,keep.lib.size=FALSE]
dge <- calcNormFactors(dge, method="TMM")

# N of genes retained after filtering
dim(dge$genes)
```

## Create design matrix
```{r}
design <- model.matrix(~0+dge$samples$gender_tissue_viral+dge$samples$library_type)
colnames(design) <- gsub("dge\\$samples\\$gender_tissue_viral", "", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$library_typeunstranded", "library_type", colnames(design))
head(design)
```

## Voom analysis
```{r}
v <- voomWithQualityWeights(dge, design, plot=TRUE)
```

# DEA
```{r}
# Block design for individual. This is used in tumor-normal comparisons with
# paired samples.
corfit <- duplicateCorrelation(v, design, block = v$targets$ID)
# This should give a positive correlation value. It represents the
# correlation between measurements made on the same person.
corfit$consensus

# Fitting the linear model with limma.
# If using paired samples, the within-patient correlation and a block design
# for patient is used to account for pairwise samples
fit <- lmFit(v, design, block = v$targets$ID, correlation = corfit$consensus)

# Contrast design for differential expression
# Defining pairwise comparisons
contrasts <- makeContrasts(F_1_HBV - F_0_HBV,
                           F_1_HCV - F_0_HCV,
                           M_1_HBV - M_0_HBV,
                           M_1_HCV - M_0_HCV,
                           levels=colnames(design))
head(contrasts)

# Running contrast analysis
vfit <- contrasts.fit(fit, contrasts = contrasts)

# Computing differential expression based on the empirical Bayes moderation of
# the standard errors towards a common value. Robust = should the estimation of
# the empirical Bayes prior parameters be robustified against outlier sample
# variances?
veBayesFit <- eBayes(vfit, robust=TRUE)
# Look at N of DEGs with adj. p <0.01 and log2FC>2
summary(decideTests(veBayesFit, adjust.method = "BH", p.value = 0.05, lfc = 2))

plotSA(veBayesFit, main = "Final model: Mean-variance trend")

## Female HBV (coef 1)
vTopTable_female_HBV <- topTable(veBayesFit, n=Inf, p.value=1, lfc=0, coef = 1)
write.csv(vTopTable_female_HBV, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/downsampled_female_HBV_tumor_tumor_adj_p1_lfc0.csv")

DEGs_female_HBV <- topTable(veBayesFit, n=Inf, p.value=0.05, lfc=2, coef=1)
DEGs_female_HBV_print <- data.frame(DEGs_female_HBV$GENEID, DEGs_female_HBV$gene_name, DEGs_female_HBV$adj.P.Val, DEGs_female_HBV$logFC)
write.csv(DEGs_female_HBV_print, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/downsampled_female_HBV_tumor_tumor_adj_p0.05_lfc2.csv")

## Female HCV (coef 2)
vTopTable_female_HCV <- topTable(veBayesFit, n=Inf, p.value=1, lfc=0, coef = 2)
write.csv(vTopTable_female_HCV, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/downsampled_female_HCV_tumor_tumor_adj_p1_lfc0.csv")

DEGs_female_HCV <- topTable(veBayesFit, n=Inf, p.value=0.05, lfc=2, coef=2)
DEGs_female_HCV_print <- data.frame(DEGs_female_HCV$GENEID, DEGs_female_HCV$gene_name, DEGs_female_HCV$adj.P.Val, DEGs_female_HCV$logFC)
write.csv(DEGs_female_HCV_print, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/downsampled_female_HCV_tumor_tumor_adj_p0.05_lfc2.csv")

## Male HBV (coef 3)
vTopTable_male_HBV <- topTable(veBayesFit, n=Inf, p.value=1, lfc=0, coef = 3)
write.csv(vTopTable_male_HBV, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/downsampled_male_HBV_tumor_tumor_adj_p1_lfc0.csv")

DEGs_male_HBV <- topTable(veBayesFit, n=Inf, p.value=0.05, lfc=2, coef=3)
DEGs_male_HBV_print <- data.frame(DEGs_male_HBV$GENEID, DEGs_male_HBV$gene_name, DEGs_male_HBV$adj.P.Val, DEGs_male_HBV$logFC)
write.csv(DEGs_male_HBV_print, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/downsampled_male_HBV_tumor_tumor_adj_p0.05_lfc2.csv")

## Male HCV (coef 4) 
vTopTable_male_HCV <- topTable(veBayesFit, n=Inf, p.value=1, lfc=0, coef = 4)
write.csv(vTopTable_male_HCV, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/downsampled_male_HCV_tumor_tumor_adj_p1_lfc0.csv")

DEGs_male_HCV <- topTable(veBayesFit, n=Inf, p.value=0.05, lfc=2, coef=4)
DEGs_male_HCV_print <- data.frame(DEGs_male_HCV$GENEID, DEGs_male_HCV$gene_name, DEGs_male_HCV$adj.P.Val, DEGs_male_HCV$logFC)
write.csv(DEGs_male_HCV_print, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/downsampled_male_HCV_tumor_tumor_adj_p0.05_lfc2.csv")
```


## Tumor vs. tumor-adjacent expression

### Create design matrix
```{r}
design <- model.matrix(~0+dge$samples$tissue_viral+dge$samples$library_type)
colnames(design) <- gsub("dge\\$samples\\$tissue_viral", "", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$library_typeunstranded", "library_type", colnames(design))
head(design)
```

### Voom analysis
```{r}
v <- voomWithQualityWeights(dge, design, plot=TRUE)
```

### DEA
```{r}
# Block design for individual. This is used in tumor-normal comparisons with
# paired samples.
corfit <- duplicateCorrelation(v, design, block = v$targets$ID)
# This should give a positive correlation value. It represents the
# correlation between measurements made on the same person.
corfit$consensus

# Fitting the linear model with limma.
# If using paired samples, the within-patient correlation and a block design
# for patient is used to account for pairwise samples
fit <- lmFit(v, design, block = v$targets$ID, correlation = corfit$consensus)

# Contrast design for differential expression
# Defining pairwise comparisons
contrasts <- makeContrasts(HBV_tumor_adjacent = HBV_1 - HBV_0,
                           HCV_tumor_ajcanet = HCV_1 - HCV_0,
                           levels=colnames(design))
head(contrasts)

# Running contrast analysis
vfit <- contrasts.fit(fit, contrasts = contrasts)

# Computing differential expression based on the empirical Bayes moderation of
# the standard errors towards a common value. Robust = should the estimation of
# the empirical Bayes prior parameters be robustified against outlier sample
# variances?
veBayesFit <- eBayes(vfit, robust=TRUE)
# Look at N of DEGs with adj. p <0.01 and log2FC>2
summary(decideTests(veBayesFit, adjust.method = "BH", p.value = 0.05, lfc = 2))

plotSA(veBayesFit, main = "Final model: Mean-variance trend")

vTopTable_HBV <- topTable(veBayesFit, n=Inf, p.value=1, lfc=0, coef = 1)
write.csv(vTopTable_HBV, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/downsampled_HBV_tumor_tumor_adj_p1_lfc0.csv")

DEGs_HBV <- topTable(veBayesFit, n=Inf, p.value=0.05, lfc=2, coef=1)
DEGs_HBV_print <- data.frame(DEGs_HBV$GENEID, DEGs_HBV$gene_name, DEGs_HBV$adj.P.Val, DEGs_HBV$logFC)
write.csv(DEGs_HBV_print, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/downsampled_HBV_tumor_tumor_adj_p0.05_lfc2.csv")

vTopTable_HCV <- topTable(veBayesFit, n=Inf, p.value=1, lfc=0, coef = 2)
write.csv(vTopTable_HCV, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/downsampled_HCV_tumor_tumor_adj_p1_lfc0.csv")

DEGs_HCV <- topTable(veBayesFit, n=Inf, p.value=0.05, lfc=2, coef=2)
DEGs_HCV_print <- data.frame(DEGs_HCV$GENEID, DEGs_HCV$gene_name, DEGs_HCV$adj.P.Val, DEGs_HCV$logFC)
write.csv(DEGs_HCV_print, "~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/downsampled_HCV_tumor_tumor_adj_p0.05_lfc2.csv")
```

**Interpretation:**

* Model fit looks good before and after voom adjustment

* Positive correlation for samples from the same patient

* DEG lists saved


