---
title: "2024_3_4_Visualize_interesting_genes"
output: html_document
date: "2024-03-4"
---

**Purpose:**

* Visualize genes with opposite logFC in HBV/HCV and M/F

**Methods:**

* Voom transformation on all genes

* Visualize boxplots of genes of interest

**Results:**

## Load libraries

```{r, message=FALSE, warning=FALSE}
library(edgeR)
library(DESeq2)
library(limma)
library(ggplot2)
library(dplyr)
```

## Read in data

```{r}
metadata <- read.table("~/3.0 Hasting Research/Liver Cancer Analysis/Japan_HCC_metadata_QC_filtered.csv", row.names=1,header=TRUE, sep=",")
tumorAdjacentExp <- read.csv("~/3.0 Hasting Research/Liver Cancer Analysis/Japan_HCC_Salmon_Expression_QC_filtered.csv", row.names = 1, header=TRUE)
colnames(tumorAdjacentExp) <- gsub("\\.", "-", colnames(tumorAdjacentExp))
identical(metadata$sampleid, colnames(tumorAdjacentExp))

table(metadata$gender_tissue_viral)
```

## Relabeling genes
```{r}
# Importing gene annotations
genes <- read.table("~/3.0 Hasting Research/Liver Cancer Analysis/gencodeTranscripts.txt", header=TRUE, sep="\t")
genes <- data.frame(genes)
tumorAdjacentExp <- tumorAdjacentExp[rownames(tumorAdjacentExp) %in% genes$GENEID ,]
genes <- genes[match(rownames(tumorAdjacentExp), genes$GENEID),]
identical(rownames(tumorAdjacentExp), genes$GENEID)
# Calculating gene length, this is needed for calculating the FPKM values
genes$length <- with(genes, end - start)
```

## Create DEG list object
```{r}
dge <- DGEList(counts=tumorAdjacentExp, genes=genes)

colnames(dge) <- colnames(tumorAdjacentExp)
dge$samples$ID <- metadata$ID
dge$samples$tumor <- factor(metadata$tumor)
dge$samples$library_type <- metadata$libtype
dge$samples$viral <- metadata$Virus_infection
dge$samples$tissue_viral <- paste(metadata$Virus_infection, metadata$tumor, sep="_")
dge$samples$gender_tussue <- metadata$gender_tissue
dge$samples$gender_tissue_viral <- metadata$gender_tissue_viral
```

## Filtering by FPKM expression level
```{r}
# Keeping genes that have a mean FPKM of at least 0.5 in at least one of the
# groups under investigation and at least 6 reads in at least 10 samples
fpkm <- rpkm(dge, gene.length=dge$genes$length)

M_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HBV")],1,mean,na.rm=TRUE)
M_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HBV")],1,mean,na.rm=TRUE)
M_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HCV")],1,mean,na.rm=TRUE)
M_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HCV")],1,mean,na.rm=TRUE)

F_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HBV")],1,mean,na.rm=TRUE)
F_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HBV")],1,mean,na.rm=TRUE)
F_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HCV")],1,mean,na.rm=TRUE)
F_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HCV")],1,mean,na.rm=TRUE)

keep <- (M_1_HBV_mean_fpkm > 0.5 | M_0_HBV_mean_fpkm > 0.5 | 
           M_1_HCV_mean_fpkm > 0.5 | M_0_HCV_mean_fpkm > 0.5 |
           F_1_HBV_mean_fpkm > 0.5 | F_0_HBV_mean_fpkm > 0.5 |
           F_1_HCV_mean_fpkm > 0.5 | F_0_HCV_mean_fpkm > 0.5 )

dge <- dge[keep,,keep.lib.sizes=FALSE]
dge <- calcNormFactors(dge, method="TMM")
keep <- rowSums(dge$counts > 6) >= 10
dge <- dge[keep,,keep.lib.size=FALSE]
dge <- calcNormFactors(dge, method="TMM")

# N of genes retained after filtering
dim(dge$genes)
```

## Create design matrix
```{r}
design <- model.matrix(~0+dge$samples$gender_tussue+dge$samples$library_type)
colnames(design) <- gsub("dge\\$samples\\$gender_tussue", "", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$library_typeunstranded", "library_type", colnames(design))
head(design)
```

## Voom analysis
```{r}
v <- voomWithQualityWeights(dge, design, plot=TRUE)
```

## Test expression pre- versus post- voom

```{r}
vexpr <- as.data.frame(v$E)
identical(colnames(vexpr), metadata$sampleid)
M_T <- vexpr[,which(metadata$tumor==1 & metadata$Gender=="M")]
M_Ta <- vexpr[,which(metadata$tumor==0 & metadata$Gender=="M")]
F_T <- vexpr[,which(metadata$tumor==1 & metadata$Gender=="F")]
F_Ta <- vexpr[,which(metadata$tumor==0 & metadata$Gender=="F")]

genes <- genes[match(row.names(M_T), genes$GENEID),]
identical(row.names(M_T), genes$GENEID)


M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="LINC01370"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="LINC01370"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="LINC01370"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="LINC01370"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000237767.1))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("LINC01370")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/LINC01370_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000237767.1))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("LINC01370")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="MEAF6"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="MEAF6"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="MEAF6"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="MEAF6"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000163875.15))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("MEAF6")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/MEAF6_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000163875.15))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("MEAF6")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="MIR6753"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="MIR6753"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="MIR6753"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="MIR6753"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000275022.1))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("MIR6753")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/MIR6753_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000275022.1))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("MIR6753")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="GPR153"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="GPR153"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="GPR153"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="GPR153"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000158292.6))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/GPR153_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000158292.6))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("GPR153")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="INO80E"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="INO80E"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="INO80E"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="INO80E"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000169592.14))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("INO80E")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/INO80E_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000169592.14))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("INO80E")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="ADGRD1"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="ADGRD1"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="ADGRD1"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="ADGRD1"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000111452.12))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("ADGRD1")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/ADGRD1_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000111452.12))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("ADGRD1")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="MRNIP"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="MRNIP"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="MRNIP"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="MRNIP"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000161010.14))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("MRNIP")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/MRNIP_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000161010.14))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("MRNIP")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="KLK4"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="KLK4"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="KLK4"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="KLK4"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000167749.11))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("KLK4")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/KLK4_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000167749.11))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("KLK4")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="ACP2"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="ACP2"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="ACP2"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="ACP2"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000134575.9))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("ACP2")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/ACP2_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000134575.9))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("ACP2")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="KCTD13"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="KCTD13"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="KCTD13"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="KCTD13"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000174943.10))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("KCTD13")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/KCTD13_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000174943.10))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("KCTD13")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="MIR483"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="MIR483"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="MIR483"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="MIR483"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000207805.3))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("MIR483")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/MIR483_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000207805.3))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("MIR483")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="TELO2"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="TELO2"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="TELO2"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="TELO2"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000100726.14))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("TELO2")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/TELO2_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000100726.14))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("TELO2")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="BET1"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="BET1"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="BET1"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="BET1"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000105829.12))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("BET1")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/BET1_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000105829.12))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("BET1")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="EVPL"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="EVPL"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="EVPL"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="EVPL"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000167880.7))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("EVPL")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/EVPL_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000167880.7))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("EVPL")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="RPL12P14"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="RPL12P14"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="RPL12P14"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="RPL12P14"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000224321.1))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("RPL12P14")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/RPL12P14_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000224321.1))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("RPL12P14")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="RABL2B"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="RABL2B"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="RABL2B"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="RABL2B"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000079974.17))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("RABL2B")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/RABL2B_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000079974.17))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("RABL2B")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="AC004233.3"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="AC004233.3"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="AC004233.3"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="AC004233.3"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000272079.2))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("AC004233.3")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/AC004233.3_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000272079.2))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("AC004233.3")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="HIRIP3"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="HIRIP3"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="HIRIP3"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="HIRIP3"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000149929.15))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("HIRIP3")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/HIRIP3_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000149929.15))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("HIRIP3")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="GNPTG"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="GNPTG"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="GNPTG"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="GNPTG"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000090581.9))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("GNPTG")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/GNPTG_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000090581.9))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("GNPTG")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="ZNF862"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="ZNF862"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="ZNF862"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="ZNF862"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000106479.10))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("ZNF862")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/ZNF862_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000106479.10))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("ZNF862")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="SLC26A1"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="SLC26A1"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="SLC26A1"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="SLC26A1"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000145217.13))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("SLC26A1")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/SLC26A1_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000145217.13))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("SLC26A1")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="SLC9A3-AS1"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="SLC9A3-AS1"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="SLC9A3-AS1"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="SLC9A3-AS1"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000225138.7))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("SLC9A3-AS1")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/SLC9A3-AS1_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000225138.7))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("SLC9A3-AS1")
dev.off()
```

```{r}
M_T_tg <- as.data.frame(t(M_T[which(genes$gene_name=="STARD8"),]))
M_T_tg$group <- "M_T"
M_Ta_tg <- as.data.frame(t(M_Ta[which(genes$gene_name=="STARD8"),]))
M_Ta_tg$group <- "M_Ta"
F_T_tg <- as.data.frame(t(F_T[which(genes$gene_name=="STARD8"),]))
F_T_tg$group <- "F_T"
F_Ta_tg <- as.data.frame(t(F_Ta[which(genes$gene_name=="STARD8"),]))
F_Ta_tg$group <- "F_ta"

target_gene <- rbind(M_T_tg, M_Ta_tg, F_T_tg, F_Ta_tg)
ggplot(target_gene, aes(x=group, y=ENSG00000130052.13))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()+ ylab("STARD8")

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/Boxplots_opposite_direction_genes/STARD8_boxplot.pdf")
ggplot(target_gene, aes(x=group, y=ENSG00000130052.13))+ geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal() + ylab("STARD8")
dev.off()
```

**Interpretation:**

* Most genes are consistent with the limma/voom pathway predictions

* A few do not seem consistent

* A few are also associated with survival in TCGA based on GEPIA2 analysis - consider including in presentation

**Future steps:**

* Integrate into a figure

* Repeat for HBV/HCV genes

