---
title: "2024_1_26_Create_visualization_DEA_across_all_comparisons"
output: html_document
date: "2024-01-26"
---

**Purpose:** Create all visualizations for the differential expression analyses done in each comparison

**Methods:**

Visualize each of the following comparisons with volcano plots, upset plots, and logFC plots

* Tumor vs. Tumor-adjacent

* Tumor vs. Tumor-adjacent sex stratified

* Tumor vs. Tumor-adjacent viral stratified

* Tumor vs. Tumor-adjacent viral and sex stratified


**Results:**


## Libraries

```{r}
library(ggrepel)
library(ggplot2)
library(UpSetR)
library(clusterProfiler)
library(org.Hs.eg.db)
```

## Read in files

```{r}
# F HBV tumor tumor-adjacent
tumor_adjacent_F_HBV_uf <- as.data.frame(read.csv("~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/female_HBV_tumor_tumor_adj_p1_lfc0.csv"))
tumor_adjacent_F_HBV <- as.data.frame(read.csv("~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/female_HBV_tumor_tumor_adj_p0.05_lfc2.csv"))
colnames(tumor_adjacent_F_HBV) <- colnames(tumor_adjacent)

# F HCV tumor tumor-adjacent
tumor_adjacent_F_HCV_uf <- as.data.frame(read.csv("~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/female_HCV_tumor_tumor_adj_p1_lfc0.csv"))
tumor_adjacent_F_HCV <- as.data.frame(read.csv("~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/female_HBV_tumor_tumor_adj_p0.05_lfc2.csv"))
colnames(tumor_adjacent_F_HCV) <- colnames(tumor_adjacent)

# M HBV tumor tumor-adjacent
tumor_adjacent_M_HBV_uf <- as.data.frame(read.csv("~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/male_HBV_tumor_tumor_adj_p1_lfc0.csv"))
tumor_adjacent_M_HBV <- as.data.frame(read.csv("~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/male_HBV_tumor_tumor_adj_p0.05_lfc2.csv"))
colnames(tumor_adjacent_M_HBV) <- colnames(tumor_adjacent)

# M HCV tumor tumor-adjacent
tumor_adjacent_M_HCV_uf <- as.data.frame(read.csv("~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/male_HCV_tumor_tumor_adj_p1_lfc0.csv"))
tumor_adjacent_M_HCV <- as.data.frame(read.csv("~/3.0 Hasting Research/Liver Cancer Analysis/DEGs/male_HCV_tumor_tumor_adj_p0.05_lfc2.csv"))
colnames(tumor_adjacent_M_HCV) <- colnames(tumor_adjacent)
```


## Print lists for females vs. males

### Significant same direction

```{r}
tumor_adjacent_M_uf <- tumor_adjacent_M_uf[match(tumor_adjacent_F_uf$GENEID, tumor_adjacent_M_uf$GENEID),]
identical(tumor_adjacent_F_uf$GENEID, tumor_adjacent_M_uf$GENEID)

tumor_adjacent_M_uf_abbr <- tumor_adjacent_M_uf[which((tumor_adjacent_M_uf$adj.P.Val<0.05 & tumor_adjacent_M_uf$logFC>0 & tumor_adjacent_F_uf$adj.P.Val<0.05 & tumor_adjacent_F_uf$logFC>0) | (tumor_adjacent_M_uf$adj.P.Val<0.05 & tumor_adjacent_M_uf$logFC<0 & tumor_adjacent_F_uf$adj.P.Val<0.05 & tumor_adjacent_F_uf$logFC<0)),]
tumor_adjacent_F_uf_abbr <- tumor_adjacent_F_uf[which((tumor_adjacent_M_uf$adj.P.Val<0.05 & tumor_adjacent_M_uf$logFC>0 & tumor_adjacent_F_uf$adj.P.Val<0.05 & tumor_adjacent_F_uf$logFC>0) | (tumor_adjacent_M_uf$adj.P.Val<0.05 & tumor_adjacent_M_uf$logFC<0 & tumor_adjacent_F_uf$adj.P.Val<0.05 & tumor_adjacent_F_uf$logFC<0)),]
identical(tumor_adjacent_F_uf_abbr$GENEID, tumor_adjacent_M_uf_abbr$GENEID)

plot(tumor_adjacent_M_uf$logFC, tumor_adjacent_F_uf$logFC, type="n")
abline(h=0, col="darkgray", lty=3)
abline(v=0, col="darkgray", lty=3)
points(tumor_adjacent_M_uf_abbr$logFC, tumor_adjacent_F_uf_abbr$logFC, 
     col=ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val>=0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val>=0.05, "black",
                ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val>0.05,"blue",
                       ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val>0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val<0.05, "maroon",
                              ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val<0.05, "goldenrod", "red")))))

write.csv(tumor_adjacent_M_uf_abbr, "~/3.0 Hasting Research/Liver Cancer Analysis/Gene lists/significant_male_female_same_dir.csv")
```

### GO/KEGG genes shared same direction male/female

```{r}
ego <- enrichGO(gene          = tumor_adjacent_M_uf_abbr$gene_name,
                OrgDb         = org.Hs.eg.db,
                keyType       = "SYMBOL",
                ont           = "BP",
                pAdjustMethod = "BH",
                qvalueCutoff  = 0.05,
        readable      = TRUE)
head(ego)
barplot(ego, showCategory=20) 

write.csv(ego, "~/3.0 Hasting Research/Liver Cancer Analysis/Pathway_results/GO_terms_significant_male_female_same_dir.csv")
```

```{r}
tumor_adjacent_M_uf <- tumor_adjacent_M_uf[match(tumor_adjacent_F_uf$GENEID, tumor_adjacent_M_uf$GENEID),]
identical(tumor_adjacent_F_uf$GENEID, tumor_adjacent_M_uf$GENEID)

tumor_adjacent_M_uf_abbr <- tumor_adjacent_M_uf[which((tumor_adjacent_M_uf$adj.P.Val<0.05 & tumor_adjacent_M_uf$logFC>2 & tumor_adjacent_F_uf$adj.P.Val<0.05 & tumor_adjacent_F_uf$logFC>2) | (tumor_adjacent_M_uf$adj.P.Val<0.05 & tumor_adjacent_M_uf$logFC< -2 & tumor_adjacent_F_uf$adj.P.Val<0.05 & tumor_adjacent_F_uf$logFC< -2)),]
tumor_adjacent_F_uf_abbr <- tumor_adjacent_F_uf[which((tumor_adjacent_M_uf$adj.P.Val<0.05 & tumor_adjacent_M_uf$logFC>2 & tumor_adjacent_F_uf$adj.P.Val<0.05 & tumor_adjacent_F_uf$logFC>2) | (tumor_adjacent_M_uf$adj.P.Val<0.05 & tumor_adjacent_M_uf$logFC< -2 & tumor_adjacent_F_uf$adj.P.Val<0.05 & tumor_adjacent_F_uf$logFC< -2)),]
identical(tumor_adjacent_F_uf_abbr$GENEID, tumor_adjacent_M_uf_abbr$GENEID)

plot(tumor_adjacent_M_uf$logFC, tumor_adjacent_F_uf$logFC, type="n")
abline(h=0, col="darkgray", lty=3)
abline(v=0, col="darkgray", lty=3)
points(tumor_adjacent_M_uf_abbr$logFC, tumor_adjacent_F_uf_abbr$logFC, 
     col=ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val>=0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val>=0.05, "black",
                ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val>0.05,"blue",
                       ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val>0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val<0.05, "maroon",
                              ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val<0.05, "goldenrod", "red")))))

#write.csv(tumor_adjacent_M_uf_abbr, "~/3.0 Hasting Research/Liver Cancer Analysis/Gene lists/significant_male_female_same_dir.csv")
```

```{r}
ego <- enrichGO(gene          = tumor_adjacent_M_uf_abbr$gene_name,
                OrgDb         = org.Hs.eg.db,
                keyType       = "SYMBOL",
                ont           = "BP",
                pAdjustMethod = "BH",
                qvalueCutoff  = 0.05,
        readable      = TRUE)
head(ego)
barplot(ego, showCategory=20) 

write.csv(ego, "~/3.0 Hasting Research/Liver Cancer Analysis/Pathway_results/GO_terms_significant_male_female_same_dir_lfc2.csv")
```

### Significant in both opposite direction

```{r}
tumor_adjacent_M_uf_abbr <- tumor_adjacent_M_uf[which((tumor_adjacent_M_uf$adj.P.Val<0.05 & tumor_adjacent_M_uf$logFC>0 & tumor_adjacent_F_uf$adj.P.Val<0.05 & tumor_adjacent_F_uf$logFC<0) | (tumor_adjacent_M_uf$adj.P.Val<0.05 & tumor_adjacent_M_uf$logFC<0 & tumor_adjacent_F_uf$adj.P.Val<0.05 & tumor_adjacent_F_uf$logFC>0)),]
tumor_adjacent_F_uf_abbr <- tumor_adjacent_F_uf[which((tumor_adjacent_M_uf$adj.P.Val<0.05 & tumor_adjacent_M_uf$logFC>0 & tumor_adjacent_F_uf$adj.P.Val<0.05 & tumor_adjacent_F_uf$logFC<0) | (tumor_adjacent_M_uf$adj.P.Val<0.05 & tumor_adjacent_M_uf$logFC<0 & tumor_adjacent_F_uf$adj.P.Val<0.05 & tumor_adjacent_F_uf$logFC>0)),]
identical(tumor_adjacent_F_uf_abbr$GENEID, tumor_adjacent_M_uf_abbr$GENEID)

plot(tumor_adjacent_M_uf_abbr$logFC, tumor_adjacent_F_uf_abbr$logFC, type="n")
abline(h=0, col="darkgray", lty=3)
abline(v=0, col="darkgray", lty=3)
points(tumor_adjacent_M_uf_abbr$logFC, tumor_adjacent_F_uf_abbr$logFC, 
     col=ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val>=0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val>=0.05, "black",
                ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val>0.05,"blue",
                       ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val>0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val<0.05, "maroon",
                              ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val<0.05, "goldenrod", "red")))))

write.csv(tumor_adjacent_M_uf_abbr, "~/3.0 Hasting Research/Liver Cancer Analysis/Gene lists/significant_male_female_opp_dir.csv")
```

### GO/KEGG genes shared opposite direction male/female

```{r}
ego <- enrichGO(gene          = tumor_adjacent_M_uf_abbr$gene_name,
                OrgDb         = org.Hs.eg.db,
                keyType       = "SYMBOL",
                ont           = "BP",
                pAdjustMethod = "BH",
                qvalueCutoff  = 0.05,
        readable      = TRUE)
#head(ego)
#barplot(ego, showCategory=20) 
```


### Significant in males only

```{r}
tumor_adjacent_M_uf_abbr <- tumor_adjacent_M_uf[which(tumor_adjacent_M_uf$adj.P.Val<0.05 & tumor_adjacent_F_uf$adj.P.Val>=0.05),]
tumor_adjacent_F_uf_abbr <- tumor_adjacent_F_uf[which(tumor_adjacent_M_uf$adj.P.Val<0.05 & tumor_adjacent_F_uf$adj.P.Val>=0.05),]
identical(tumor_adjacent_F_uf_abbr$GENEID, tumor_adjacent_M_uf_abbr$GENEID)

plot(tumor_adjacent_M_uf_abbr$logFC, tumor_adjacent_F_uf_abbr$logFC, type="n")
abline(h=0, col="darkgray", lty=3)
abline(v=0, col="darkgray", lty=3)
points(tumor_adjacent_M_uf_abbr$logFC, tumor_adjacent_F_uf_abbr$logFC, 
     col=ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val>=0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val>=0.05, "black",
                ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val>0.05,"blue",
                       ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val>0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val<0.05, "maroon",
                              ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val<0.05, "goldenrod", "red")))))
```

### GO/KEGG genes significant male only

```{r}
ego <- enrichGO(gene          = tumor_adjacent_M_uf_abbr$gene_name,
                OrgDb         = org.Hs.eg.db,
                keyType       = "SYMBOL",
                ont           = "BP",
                pAdjustMethod = "BH",
                qvalueCutoff  = 0.05,
        readable      = TRUE)
head(ego)
barplot(ego, showCategory=20) 

write.csv(ego, "~/3.0 Hasting Research/Liver Cancer Analysis/Pathway_results/GO_terms_significant_male_only.csv")
```

```{r}
tumor_adjacent_M_uf_abbr <- tumor_adjacent_M_uf[which(tumor_adjacent_M_uf$adj.P.Val>=0.05 & tumor_adjacent_F_uf$adj.P.Val<0.05),]
tumor_adjacent_F_uf_abbr <- tumor_adjacent_F_uf[which(tumor_adjacent_M_uf$adj.P.Val>=0.05 & tumor_adjacent_F_uf$adj.P.Val<0.05),]
identical(tumor_adjacent_F_uf_abbr$GENEID, tumor_adjacent_M_uf_abbr$GENEID)

plot(tumor_adjacent_M_uf_abbr$logFC, tumor_adjacent_F_uf_abbr$logFC, type="n")
abline(h=0, col="darkgray", lty=3)
abline(v=0, col="darkgray", lty=3)
points(tumor_adjacent_M_uf_abbr$logFC, tumor_adjacent_F_uf_abbr$logFC, 
     col=ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val>=0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val>=0.05, "black",
                ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val>0.05,"blue",
                       ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val>0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val<0.05, "maroon",
                              ifelse(tumor_adjacent_M_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_F_uf_abbr$adj.P.Val<0.05, "goldenrod", "red")))))
```

### GO/KEGG genes significant female only

```{r}
ego <- enrichGO(gene          = tumor_adjacent_M_uf_abbr$gene_name,
                OrgDb         = org.Hs.eg.db,
                keyType       = "SYMBOL",
                ont           = "BP",
                pAdjustMethod = "BH",
                qvalueCutoff  = 0.05,
        readable      = TRUE)
head(ego)
barplot(ego, showCategory=20)
write.csv(ego, "~/3.0 Hasting Research/Liver Cancer Analysis/Pathway_results/GO_terms_significant_female_only.csv")
```

## Print lists for HBV vs. HCV

### Significant same direction

```{r}
tumor_adjacent_HBV_uf <- tumor_adjacent_HBV_uf[match(tumor_adjacent_HCV_uf$GENEID, tumor_adjacent_HBV_uf$GENEID),]
identical(tumor_adjacent_HCV_uf$GENEID, tumor_adjacent_HBV_uf$GENEID)

tumor_adjacent_HBV_uf_abbr <- tumor_adjacent_HBV_uf[which((tumor_adjacent_HBV_uf$adj.P.Val<0.05 & tumor_adjacent_HBV_uf$logFC> 2 & tumor_adjacent_HCV_uf$adj.P.Val<0.05 & tumor_adjacent_HCV_uf$logFC> 2) | (tumor_adjacent_HBV_uf$adj.P.Val<0.05 & tumor_adjacent_HBV_uf$logFC< -2 & tumor_adjacent_HCV_uf$adj.P.Val<0.05 & tumor_adjacent_HCV_uf$logFC< -2)),]
tumor_adjacent_HCV_uf_abbr <- tumor_adjacent_HCV_uf[which((tumor_adjacent_HBV_uf$adj.P.Val<0.05 & tumor_adjacent_HBV_uf$logFC>2 & tumor_adjacent_HCV_uf$adj.P.Val<0.05 & tumor_adjacent_HCV_uf$logFC>2) | (tumor_adjacent_HBV_uf$adj.P.Val<0.05 & tumor_adjacent_HBV_uf$logFC< -2 & tumor_adjacent_HCV_uf$adj.P.Val<0.05 & tumor_adjacent_HCV_uf$logFC< -2)),]
identical(tumor_adjacent_HCV_uf_abbr$GENEID, tumor_adjacent_HBV_uf_abbr$GENEID)

plot(tumor_adjacent_HBV_uf$logFC, tumor_adjacent_HCV_uf$logFC, type="n")
abline(h=0, col="darkgray", lty=3)
abline(v=0, col="darkgray", lty=3)
points(tumor_adjacent_HBV_uf_abbr$logFC, tumor_adjacent_HCV_uf_abbr$logFC, 
     col=ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val>=0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val>=0.05, "black",
                ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val>0.05,"blue",
                       ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val>0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val<0.05, "maroon",
                              ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val<0.05, "goldenrod", "red")))))
```

### GO/KEGG genes shared same direction HBV/HCV

```{r}
ego <- enrichGO(gene          = tumor_adjacent_HBV_uf_abbr$gene_name,
                OrgDb         = org.Hs.eg.db,
                keyType       = "SYMBOL",
                ont           = "BP",
                pAdjustMethod = "BH",
                qvalueCutoff  = 0.05,
        readable      = TRUE)
head(ego)
barplot(ego, showCategory=20) 

write.csv(ego, "~/3.0 Hasting Research/Liver Cancer Analysis/Pathway_results/GO_terms_significant_HBV_HCV_same_dir_lfc2.csv")
```

### Significant in both opposite direction

```{r}
tumor_adjacent_HBV_uf_abbr <- tumor_adjacent_HBV_uf[which((tumor_adjacent_HBV_uf$adj.P.Val<0.05 & tumor_adjacent_HBV_uf$logFC>0 & tumor_adjacent_HCV_uf$adj.P.Val<0.05 & tumor_adjacent_HCV_uf$logFC<0) | (tumor_adjacent_HBV_uf$adj.P.Val<0.05 & tumor_adjacent_HBV_uf$logFC<0 & tumor_adjacent_HCV_uf$adj.P.Val<0.05 & tumor_adjacent_HCV_uf$logFC>0)),]
tumor_adjacent_HCV_uf_abbr <- tumor_adjacent_HCV_uf[which((tumor_adjacent_HBV_uf$adj.P.Val<0.05 & tumor_adjacent_HBV_uf$logFC>0 & tumor_adjacent_HCV_uf$adj.P.Val<0.05 & tumor_adjacent_HCV_uf$logFC<0) | (tumor_adjacent_HBV_uf$adj.P.Val<0.05 & tumor_adjacent_HBV_uf$logFC<0 & tumor_adjacent_HCV_uf$adj.P.Val<0.05 & tumor_adjacent_HCV_uf$logFC>0)),]
identical(tumor_adjacent_HCV_uf_abbr$GENEID, tumor_adjacent_HBV_uf_abbr$GENEID)

plot(tumor_adjacent_HBV_uf_abbr$logFC, tumor_adjacent_HCV_uf_abbr$logFC, type="n")
abline(h=0, col="darkgray", lty=3)
abline(v=0, col="darkgray", lty=3)
points(tumor_adjacent_HBV_uf_abbr$logFC, tumor_adjacent_HCV_uf_abbr$logFC, 
     col=ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val>=0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val>=0.05, "black",
                ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val>0.05,"blue",
                       ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val>0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val<0.05, "maroon",
                              ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val<0.05, "goldenrod", "red")))))
```

### GO/KEGG genes shared opposite direction HBV/HCV

```{r}
ego <- enrichGO(gene          = tumor_adjacent_HBV_uf_abbr$gene_name,
                OrgDb         = org.Hs.eg.db,
                keyType       = "SYMBOL",
                ont           = "BP",
                pAdjustMethod = "BH",
                qvalueCutoff  = 0.05,
        readable      = TRUE)
head(ego)
barplot(ego, showCategory=20) 
write.csv(ego, "~/3.0 Hasting Research/Liver Cancer Analysis/Pathway_results/GO_terms_significant_HBV_HCV_opposite_directions.csv")
```

### Significant in HBV

```{r}
tumor_adjacent_HBV_uf_abbr <- tumor_adjacent_HBV_uf[which(tumor_adjacent_HBV_uf$adj.P.Val<0.05 & tumor_adjacent_HCV_uf$adj.P.Val>=0.05),]
tumor_adjacent_HCV_uf_abbr <- tumor_adjacent_HCV_uf[which(tumor_adjacent_HBV_uf$adj.P.Val<0.05 & tumor_adjacent_HCV_uf$adj.P.Val>=0.05),]
identical(tumor_adjacent_HCV_uf_abbr$GENEID, tumor_adjacent_HBV_uf_abbr$GENEID)

plot(tumor_adjacent_HBV_uf_abbr$logFC, tumor_adjacent_HCV_uf_abbr$logFC, type="n")
abline(h=0, col="darkgray", lty=3)
abline(v=0, col="darkgray", lty=3)
points(tumor_adjacent_HBV_uf_abbr$logFC, tumor_adjacent_HCV_uf_abbr$logFC, 
     col=ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val>=0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val>=0.05, "black",
                ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val>0.05,"blue",
                       ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val>0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val<0.05, "maroon",
                              ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val<0.05, "goldenrod", "red")))))
```

### GO/KEGG genes significant HBV only

```{r}
ego <- enrichGO(gene          = tumor_adjacent_HBV_uf_abbr$gene_name,
                OrgDb         = org.Hs.eg.db,
                keyType       = "SYMBOL",
                ont           = "BP",
                pAdjustMethod = "BH",
                qvalueCutoff  = 0.05,
        readable      = TRUE)
head(ego)
barplot(ego, showCategory=20) 
write.csv(ego, "~/3.0 Hasting Research/Liver Cancer Analysis/Pathway_results/GO_terms_significant_HBV_only.csv")
```

```{r}
tumor_adjacent_HBV_uf_abbr <- tumor_adjacent_HBV_uf[which(tumor_adjacent_HBV_uf$adj.P.Val>=0.05 & tumor_adjacent_HCV_uf$adj.P.Val<0.05),]
tumor_adjacent_HCV_uf_abbr <- tumor_adjacent_HCV_uf[which(tumor_adjacent_HBV_uf$adj.P.Val>=0.05 & tumor_adjacent_HCV_uf$adj.P.Val<0.05),]
identical(tumor_adjacent_HCV_uf_abbr$GENEID, tumor_adjacent_HBV_uf_abbr$GENEID)

plot(tumor_adjacent_HBV_uf_abbr$logFC, tumor_adjacent_HCV_uf_abbr$logFC, type="n")
abline(h=0, col="darkgray", lty=3)
abline(v=0, col="darkgray", lty=3)
points(tumor_adjacent_HBV_uf_abbr$logFC, tumor_adjacent_HCV_uf_abbr$logFC, 
     col=ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val>=0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val>=0.05, "black",
                ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val>0.05,"blue",
                       ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val>0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val<0.05, "maroon",
                              ifelse(tumor_adjacent_HBV_uf_abbr$adj.P.Val<0.05 & tumor_adjacent_HCV_uf_abbr$adj.P.Val<0.05, "goldenrod", "red")))))
```

### GO/KEGG genes significant HCV only

```{r}
ego <- enrichGO(gene          = tumor_adjacent_HBV_uf_abbr$gene_name,
                OrgDb         = org.Hs.eg.db,
                keyType       = "SYMBOL",
                ont           = "BP",
                pAdjustMethod = "BH",
                qvalueCutoff  = 0.05,
        readable      = TRUE)
head(ego)
barplot(ego, showCategory=20)
write.csv(ego, "~/3.0 Hasting Research/Liver Cancer Analysis/Pathway_results/GO_terms_significant_HCV_only.csv")
```

**Interpretation:**

* Gene lists generated for all comparisons except sex and viral differentiated

**Future steps:**

* Visualize GO terms with Revigo

* Finish comparisons separated by sex and viral status
