---
title: "2024_1_22_DEA_Tumor_TumorAdj"
output: html_document
date: "2024-01-22"
---

**Purpose:**

* Identify tumor/tumor-adjacent differentially expressed genes with only library type as a covariate

**Methods:**

* Limma voom pipeline on all tumor-tumor-adjcent comparison

* Only covariate is library type

**Results:**

## Load libraries

```{r, message=FALSE, warning=FALSE}
library(ggrepel)
library(UpSetR)
library(edgeR)
library(DESeq2)
library(limma)
library(ggplot2)
library(dplyr)
library(clusterProfiler)
library(org.Hs.eg.db)
```

## Read in data

```{r}
metadata <- read.table("~/3.0 Hasting Research/Liver Cancer Analysis/Japan_HCC_metadata_QC_filtered.csv", row.names=1,header=TRUE, sep=",")
tumorAdjacentExp <- read.csv("~/3.0 Hasting Research/Liver Cancer Analysis/Japan_HCC_Salmon_Expression_QC_filtered.csv", row.names = 1, header=TRUE)
colnames(tumorAdjacentExp) <- gsub("\\.", "-", colnames(tumorAdjacentExp))
identical(metadata$sampleid, colnames(tumorAdjacentExp))

table(metadata$gender_tissue_viral)
```

## Relabeling genes
```{r}
# Importing gene annotations
genes <- read.table("~/3.0 Hasting Research/Liver Cancer Analysis/gencodeTranscripts.txt", header=TRUE, sep="\t")
genes <- data.frame(genes)
tumorAdjacentExp <- tumorAdjacentExp[rownames(tumorAdjacentExp) %in% genes$GENEID ,]
genes <- genes[match(rownames(tumorAdjacentExp), genes$GENEID),]
identical(rownames(tumorAdjacentExp), genes$GENEID)
# Calculating gene length, this is needed for calculating the FPKM values
genes$length <- with(genes, end - start)
```

## Create DEG list object
```{r}
dge <- DGEList(counts=tumorAdjacentExp, genes=genes)

colnames(dge) <- colnames(tumorAdjacentExp)
dge$samples$ID <- metadata$ID
dge$samples$tumor <- factor(metadata$tumor)
dge$samples$library_type <- metadata$libtype
dge$samples$viral <- metadata$Virus_infection
dge$samples$viral_tissue <- paste(metadata$Virus_infection, metadata$tumor, sep="_")
dge$samples$gender_tussue <- metadata$gender_tissue
dge$samples$gender_tissue_viral <- metadata$gender_tissue_viral
```

## Filtering by FPKM expression level
```{r}
# Keeping genes that have a mean FPKM of at least 0.5 in at least one of the
# groups under investigation and at least 6 reads in at least 10 samples
fpkm <- rpkm(dge, gene.length=dge$genes$length)

M_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HBV")],1,mean,na.rm=TRUE)
M_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HBV")],1,mean,na.rm=TRUE)
M_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HCV")],1,mean,na.rm=TRUE)
M_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HCV")],1,mean,na.rm=TRUE)

F_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HBV")],1,mean,na.rm=TRUE)
F_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HBV")],1,mean,na.rm=TRUE)
F_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HCV")],1,mean,na.rm=TRUE)
F_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HCV")],1,mean,na.rm=TRUE)

keep <- (M_1_HBV_mean_fpkm > 0.5 | M_0_HBV_mean_fpkm > 0.5 | 
           M_1_HCV_mean_fpkm > 0.5 | M_0_HCV_mean_fpkm > 0.5 |
           F_1_HBV_mean_fpkm > 0.5 | F_0_HBV_mean_fpkm > 0.5 |
           F_1_HCV_mean_fpkm > 0.5 | F_0_HCV_mean_fpkm > 0.5 )

dge <- dge[keep,,keep.lib.sizes=FALSE]
dge <- calcNormFactors(dge, method="TMM")
keep <- rowSums(dge$counts > 6) >= 10
dge <- dge[keep,,keep.lib.size=FALSE]
dge <- calcNormFactors(dge, method="TMM")

# N of genes retained after filtering
dim(dge$genes)
```

## Create design matrix
```{r}
design <- model.matrix(~0+dge$samples$viral_tissue+dge$samples$library_type)
colnames(design) <- c("HBV_0", "HBV_1", "HCV_0", "HCV_1", "library_type")
head(design)
```

## Voom analysis
```{r}
v <- voomWithQualityWeights(dge, design, plot=TRUE)
```

# DEA
```{r}
# Block design for individual. This is used in tumor-normal comparisons with
# paired samples.
corfit <- duplicateCorrelation(v, design, block = v$targets$ID)
# This should give a positive correlation value. It represents the
# correlation between measurements made on the same person.
corfit$consensus

# Fitting the linear model with limma.
# If using paired samples, the within-patient correlation and a block design
# for patient is used to account for pairwise samples
fit <- lmFit(v, design, block = v$targets$ID, correlation = corfit$consensus)

# Contrast design for differential expression
# Defining pairwise comparisons
contrasts <- makeContrasts(HBV_v_HCV_adj = `HBV_0` - `HCV_0`,
                           HBV_v_HCV_tumor = `HBV_1` - `HCV_1`,
                           levels=colnames(design))
head(contrasts)

# Running contrast analysis
vfit <- contrasts.fit(fit, contrasts = contrasts)


# Computing differential expression based on the empirical Bayes moderation of
# the standard errors towards a common value. Robust = should the estimation of
# the empirical Bayes prior parameters be robustified against outlier sample
# variances?
veBayesFit <- eBayes(vfit, robust=TRUE)

plotSA(veBayesFit, main = "Final model: Mean-variance trend")
# Look at N of DEGs with adj. p <0.01 and log2FC>2
summary(decideTests(veBayesFit, adjust.method = "BH", p.value = 0.05))

vTopTable <- topTable(veBayesFit, n=Inf, p.value=1, lfc=0)

All_tumor_adjacent <- topTable(veBayesFit, coef = 1, n=Inf)
DEGs_tumor_adjacent <- topTable(veBayesFit, coef = 1, n=Inf, p.value=0.05, lfc=2)
write.csv(DEGs_tumor_adjacent, "~/3.0 Hasting Research/Liver Cancer Analysis/Gene lists/HBV_HCV_tumor_adjacent_p0.05_lfc2.csv")
DEGs_tumor_adjacent_rel <- topTable(veBayesFit, coef = 1, n=Inf, p.value=0.05, lfc=1)
DEGs_tumor_adjacent_rel2 <- topTable(veBayesFit, coef = 1, n=Inf, p.value=0.05)

All_tumor <- topTable(veBayesFit, coef = 2, n=Inf)
DEGs_tumor <- topTable(veBayesFit, coef = 2, n=Inf, p.value=0.05, lfc=2)
write.csv(DEGs_tumor, "~/3.0 Hasting Research/Liver Cancer Analysis/Gene lists/HBV_HCV_tumor_p0.05_lfc2.csv")
DEGs_tumor_rel <- topTable(veBayesFit, coef = 2, n=Inf, p.value=0.05, lfc=1)
DEGs_tumor_rel2 <- topTable(veBayesFit, coef = 2, n=Inf, p.value=0.05)
```

## Violin plots
```{r}
All_tumor_adjacent$color <- NA
## Autosomal non-significant
All_tumor_adjacent[which(!(All_tumor_adjacent$GENEID %in% DEGs_tumor_adjacent$GENEID) & All_tumor_adjacent$chr != "chrX" & All_tumor_adjacent$chr != "chrY"), 14] <- 1
## X-linked non-signifiant
All_tumor_adjacent[which(!(All_tumor_adjacent$GENEID %in% DEGs_tumor_adjacent$GENEID) & All_tumor_adjacent$chr == "chrX"), 14] <- 2
## Y-linked non-significant
All_tumor_adjacent[which(!(All_tumor_adjacent$GENEID %in% DEGs_tumor_adjacent$GENEID) & All_tumor_adjacent$chr == "chrY"), 14] <- 3
## Autosomal significant
All_tumor_adjacent[which((All_tumor_adjacent$GENEID %in% DEGs_tumor_adjacent$GENEID) & All_tumor_adjacent$chr != "chrX" & All_tumor_adjacent$chr != "chrY"), 14] <- 4
## X-linked signifiant
All_tumor_adjacent[which((All_tumor_adjacent$GENEID %in% DEGs_tumor_adjacent$GENEID) & All_tumor_adjacent$chr == "chrX"), 14] <- 5
## Y-linked significant
All_tumor_adjacent[which((All_tumor_adjacent$GENEID %in% DEGs_tumor_adjacent$GENEID) & All_tumor_adjacent$chr == "chrY"), 14] <- 6

All_tumor_adjacent$color <- as.factor(All_tumor_adjacent$color)

p <- ggplot(data = All_tumor_adjacent, aes(x = logFC, y = -log10(adj.P.Val), color=color )) +
  geom_point(alpha = 0.5, size = 5) +
  theme_bw() +
  theme(legend.position = "none") +
  xlim(c(-5, 5)) + ylim(c(0, 15)) +
  scale_color_manual(values = c("azure3", "pink", "seagreen2", "black", "mediumvioletred", "springgreen")) +
  labs(x=expression(log[2](FC)),
       y=expression(-log[10] ~ "(FDR-adjusted " ~ italic("p") ~ "-value)")) +
  theme(axis.title.x=element_text(size=12), 
        axis.text.x=element_text(size=10)) +
  theme(axis.title.y=element_text(size=12),
        axis.text.y=element_text(size=10))

forLabel <- subset(All_tumor_adjacent, adj.P.Val<=0.01 & abs(logFC)>=2)

# Adding lines for significance thresholds
p + geom_hline(yintercept = 2, colour="#000000", linetype="dashed"
) + geom_vline(xintercept = 2, colour="#000000", linetype="dashed"
) + geom_vline(xintercept = -2, colour="#000000", linetype="dashed"
) + geom_text_repel(data=forLabel, max.iter=1000, box.padding = 0.25, force=1, aes(x = logFC, y = -log10(adj.P.Val), label=gene_name, color=color), size=9)

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/DEG Figures/Tumor_adjacent_HBV_v_HCV_volcano.pdf", width=12, height=12)
p + geom_hline(yintercept = 2, colour="#000000", linetype="dashed"
) + geom_vline(xintercept = 2, colour="#000000", linetype="dashed"
) + geom_vline(xintercept = -2, colour="#000000", linetype="dashed"
) + geom_text_repel(data=forLabel, max.iter=1000, box.padding = 0.25, force=1, aes(x = logFC, y = -log10(adj.P.Val), label=gene_name, color=color), size=9)
dev.off()
```

## Violin plots
```{r}
All_tumor$color <- NA
## Autosomal non-significant
All_tumor[which(!(All_tumor$GENEID %in% DEGs_tumor$GENEID) & All_tumor$chr != "chrX" & All_tumor$chr != "chrY"), 14] <- 1
## X-linked non-signifiant
All_tumor[which(!(All_tumor$GENEID %in% DEGs_tumor$GENEID) & All_tumor$chr == "chrX"), 14] <- 2
## Y-linked non-significant
All_tumor[which(!(All_tumor$GENEID %in% DEGs_tumor$GENEID) & All_tumor$chr == "chrY"), 14] <- 3
## Autosomal significant
All_tumor[which((All_tumor$GENEID %in% DEGs_tumor$GENEID) & All_tumor$chr != "chrX" & All_tumor$chr != "chrY"), 14] <- 4
## X-linked signifiant
All_tumor[which((All_tumor$GENEID %in% DEGs_tumor$GENEID) & All_tumor$chr == "chrX"), 14] <- 5
## Y-linked significant
All_tumor[which((All_tumor$GENEID %in% DEGs_tumor$GENEID) & All_tumor$chr == "chrY"), 14] <- 6

All_tumor$color <- as.factor(All_tumor$color)

p <- ggplot(data = All_tumor, aes(x = logFC, y = -log10(adj.P.Val), color=color )) +
  geom_point(alpha = 0.5, size = 5) +
  theme_bw() +
  theme(legend.position = "none") +
  xlim(c(-5, 5)) + ylim(c(0, 15)) +
  scale_color_manual(values = c("azure3", "pink", "seagreen2", "black", "mediumvioletred", "springgreen")) +
  labs(x=expression(log[2](FC)),
       y=expression(-log[10] ~ "(FDR-adjusted " ~ italic("p") ~ "-value)")) +
  theme(axis.title.x=element_text(size=12), 
        axis.text.x=element_text(size=10)) +
  theme(axis.title.y=element_text(size=12),
        axis.text.y=element_text(size=10))

forLabel <- subset(All_tumor, adj.P.Val<=0.01 & abs(logFC)>=2)

# Adding lines for significance thresholds
p + geom_hline(yintercept = 2, colour="#000000", linetype="dashed"
) + geom_vline(xintercept = 2, colour="#000000", linetype="dashed"
) + geom_vline(xintercept = -2, colour="#000000", linetype="dashed"
) + geom_text_repel(data=forLabel, max.iter=1000, box.padding = 0.25, force=1, aes(x = logFC, y = -log10(adj.P.Val), label=gene_name, color=color), size=9)

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/DEG Figures/Tumor_HBV_v_HCV_volcano.pdf", width=12, height=12)
p + geom_hline(yintercept = 2, colour="#000000", linetype="dashed"
) + geom_vline(xintercept = 2, colour="#000000", linetype="dashed"
) + geom_vline(xintercept = -2, colour="#000000", linetype="dashed"
) + geom_text_repel(data=forLabel, max.iter=1000, box.padding = 0.25, force=1, aes(x = logFC, y = -log10(adj.P.Val), label=gene_name, color=color), size=9)
dev.off()
```
## Upset plot

```{r}
upset_list <- list("Up HBV tumor"= DEGs_tumor_rel[which(DEGs_tumor_rel$logFC>0),5], 
                   "Up HCV tumor"=DEGs_tumor_rel[which(DEGs_tumor_rel$logFC<0),5],
                   "Up HBV tumor-adjacent"=DEGs_tumor_adjacent_rel[which( DEGs_tumor_adjacent_rel$logFC>0),5],
                   "Up HCV tumor-adjacent"=DEGs_tumor_adjacent_rel[which( DEGs_tumor_adjacent_rel$logFC<0),5])
upset(fromList(upset_list), order.by = "freq", sets=c("Up HBV tumor", "Up HCV tumor", "Up HBV tumor-adjacent", "Up HCV tumor-adjacent"), nsets=6, keep.order = TRUE)

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/DEG Figures/HBV_v_HCV_DEA_Upset.pdf", width=10, height=10)
upset(fromList(upset_list), order.by = "freq", sets=c("Up HBV tumor", "Up HCV tumor", "Up HBV tumor-adjacent", "Up HCV tumor-adjacent"), nsets=6, keep.order = TRUE)
dev.off()
```

## Upset plot

```{r}
upset_list <- list("Up HBV tumor"= DEGs_tumor_rel2[which(DEGs_tumor_rel2$logFC>0),5], 
                   "Up HCV tumor"=DEGs_tumor_rel2[which(DEGs_tumor_rel2$logFC<0),5],
                   "Up HBV tumor-adjacent"=DEGs_tumor_adjacent_rel2[which( DEGs_tumor_adjacent_rel2$logFC>0),5],
                   "Up HCV tumor-adjacent"=DEGs_tumor_adjacent_rel2[which( DEGs_tumor_adjacent_rel2$logFC<0),5])
upset(fromList(upset_list), order.by = "freq", sets=c("Up HBV tumor", "Up HCV tumor", "Up HBV tumor-adjacent", "Up HCV tumor-adjacent"), nsets=6, keep.order = TRUE)

pdf("~/3.0 Hasting Research/Liver Cancer Analysis/DEG Figures/HBV_v_HCV_DEA_Upset_no_filter.pdf", width=10, height=10)
upset(fromList(upset_list), order.by = "freq", sets=c("Up HBV tumor", "Up HCV tumor", "Up HBV tumor-adjacent", "Up HCV tumor-adjacent"), nsets=6, keep.order = TRUE)
dev.off()
```

## Pathway enrichment analysis
```{r}
ego <- enrichGO(gene          = DEGs_tumor_rel2$gene_name,
                OrgDb         = org.Hs.eg.db,
                keyType       = "SYMBOL",
                ont           = "BP",
                pAdjustMethod = "BH",
                qvalueCutoff  = 0.05,
        readable      = TRUE)
head(ego)
write.csv(ego, "~/3.0 Hasting Research/Liver Cancer Analysis/Pathway_results/All_HBV_HCV_tumor_pathways.csv")

ego <- enrichGO(gene          = DEGs_tumor_adjacent_rel2$gene_name,
                OrgDb         = org.Hs.eg.db,
                keyType       = "SYMBOL",
                ont           = "BP",
                pAdjustMethod = "BH",
                qvalueCutoff  = 0.05,
        readable      = TRUE)
head(ego)
write.csv(ego, "~/3.0 Hasting Research/Liver Cancer Analysis/Pathway_results/All_HBV_HCV_tumor_adjacent_pathways.csv")
```

## Plot example gene

```{r}
exp_data <- v$E

# Tumor-adjacent only, up HCV
csag3 <- exp_data[which(row.names(exp_data)=="ENSG00000268916.6"),]
csag3 <- as.data.frame(csag3)
identical(metadata$sampleid, rownames(csag3))

metadata$csag3 <- csag3$csag3
metadata$gender_viral_tissue <- paste(metadata$Gender, metadata$Virus_infection, metadata$tumor, sep="")

ggplot(metadata, aes(x=gender_viral_tissue, y=csag3)) + geom_boxplot(outlier.shape = NULL) + geom_jitter(width=0.05) + theme_minimal()

# Tumor-adjacent only, up HCV
page4 <- exp_data[which(row.names(exp_data)=="ENSG00000101951.16"),]
page4 <- as.data.frame(page4)
identical(metadata$sampleid, rownames(page4))

metadata$page4 <- page4$page4

ggplot(metadata, aes(x=gender_viral_tissue, y=page4)) + geom_boxplot(outlier.shape = NULL) + geom_jitter(width=0.05) + theme_minimal()

# Tumor-adjacent only, up HCV
bex2 <- exp_data[which(row.names(exp_data)=="ENSG00000133134.11"),]
bex2 <- as.data.frame(bex2)
identical(metadata$sampleid, rownames(bex2))

metadata$bex2 <- bex2$bex2

ggplot(metadata, aes(x=gender_viral_tissue, y=bex2)) + geom_boxplot(outlier.shape = NULL) + geom_jitter(width=0.05) + theme_minimal()
```


**Interpretation:**

* Model fit looks good before and after voom adjustment

* Positive correlation for samples from the same patient

* DEG lists saved


